---
title: "GSS challange"
author: "Logan DeHay"
date: "10/9/2021"
output: html_document
---







```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS2018.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,race,cappun)
mydata <- na.omit(mydata)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is attitude toward capitol punishment (cappun) independent of race (race)?
If not, then whose view differs from whom??

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->



## Now with less data (more variables)

```{r}
dd2 <- mydata %>% group_by(race,cappun) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=race,y=count,fill=cappun))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=race,y=prcnt*100,fill=cappun)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
Using a bar graph of percentages for and against capital punishment, we could estimate that around 68% of the white race approved of capital punishment and 32% opposing it. Around 46% of the black race approved of capital punishment while the majority of 54% opposed it. Finally, around 56% of the other race approved of capital punishment with 44% disapproving of the law. Visually, the bar graph suggests that there is a difference between all of the groups.
  
### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->




## Now with less data

```{r}
table2 <- xtabs(~race + cappun, data=mydata)
rowPerc(table2)
colPerc(table2)
```
  
The numerical results display that 67.98% of whites favor capital punishment while 53.71% of blacks oppose it. 55.87% of other races favor capital punishment and 44.13% oppose it. Both other races and white people majorly favor this law while black people largely oppose it. Yet, there is a large difference between all of the races in terms of the solidity of their majority opinion.
  
## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

`


## Less data version

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
#fisherXY(mydata$race,mydata$cappun)
```
  
Our counts vs expected counts by the null display a discrepancy in the black and other races while the white data is very close to the expected counts by the null hypothesis. The chi squared test above suggests that blacks have the largest contributions to the chi-square statistics. However, both white and black populations contributed a large amount. In contrast, the "other" race contributed a drastically smaller amount to the chi-square statistics. These contributions can be compared above under "Contributions to the chi-square statistics." The degrees of freedom calculated of 2 is also 32 times smaller than our Chi-square statistics of 64.4 which strongly indicates a difference between groups.
  
```{r}
fisherXY = function(x, y, confL = 0, printAll=FALSE){
  # Inputs: 
  #   x: factor variable, 2 levels
  #   y: factor variable, n levels
  # confL: confidence level, default = 0.95
  #        print only those tables, results if p-value < (1-confL)
  # printAll: overide the 'print only if p-value < (1-confL), and
  #           print all tables and fisher test results
  #
  # Constraint:
  #   x and y must be same length
  #
  # Outputs:
  #   All 2x2 Fisher exact tests
  
  if(confL<0 || confL> 1){
    print('error: confL must be between 0 and 1')
    return(999)
  }
  
  if(!is.factor(x)) x=as.factor(x)
  if(!is.factor(y)) y=as.factor(y)
  
  fisherTable = as.matrix(table(x, y))
  cat('\ncontingency table:\n')
  print(fisherTable)
  cat('\n***All 2X2 tests****\n\n')

  # all 2x2 tests
  levelsY = levels(y)
  levelsX = levels(x)
  
  m = length(levelsX)
  n = length(levelsY)
  
  confLevel = 0.95
  if (confL !=0 ) confLevel=confL
  
  for (i in 1:(m-1))
       for(j in ((i+1):m))
          for (k in 1:(n-1))
             for (l in ((k+1):n)) {
                
                 fisherIndexX = c(i,j)
                 fisherIndexY = c(k,l)
                 
                 fTempTable = as.matrix(fisherTable[fisherIndexX,
                                        fisherIndexY])
                 
                 fisherTest=fisher.test(fTempTable, conf.int = TRUE, 
                              conf.level = confLevel)
                 
                 if((fisherTest$p.value < (1-confLevel) ) || printAll){
                   
                       print(c(levelsX[i], levelsX[j]))
                       print(c(levelsY[k], levelsY[l]))
                       print(fTempTable)
                       print(fisherTest)
                       cat('\n')
                 }
                 
        }
}

```

```{r}
fisherXY(mydata$cappun,mydata$race)
```
  
All of the races were included in this contingency tables and all of our p-values calculated were small enough to negate our null and explicate that there is a substantial difference between all three races on this subject. Despite them all being different, the p-value between white and black was much smaller. The p-value for black and white was 0.000000000000575 while the p-values for white~other and black~other were 0.0002819 and 0.02479. The exaggeratedly small p-value for white~black indicates that the largest difference in opinions on capital punishment was between the white and black population, followed by the black and other population, then the white and other populations.
  
# Discussion and Conclusion
  
<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
  
The master branch showed distinctions between all of the groups involved in the data, and an extreme difference between black and white populations. However, the number2 branch data illustrated no difference between white and other races. This variance between the two is derived from having a larger set of data, which causes differences to become more apparent.
  